
\section{Experiments}

\subsection{Setup}

The experiments use the GIAANN prototype with POS-driven concept columns and SANI segments. Training is executed (\texttt{useInference=False}) only with a single epoch over a fixed number of sequences.

\paragraph{Dataset.}
The corpora are a) Wikipedia from the Hugging Face dataset \texttt{wikipedia} (config \texttt{20220301.en}), and OSCAR-2201 from the Hugging Face dataset \texttt{oscar-corpus/OSCAR-2201} (config \texttt{en}). For Wikipedia, a local dataset cache is used (\texttt{useLocalDataset=True}) with manual download into \texttt{../../dataset/}, and the processed cache stored under \path{processed_dataset_cache}.

\paragraph{Preprocessing and POS handling.}
Tokens are produced by spaCy (models \texttt{en\_core\_web\_trf} or \texttt{en\_core\_web\_sm} with \texttt{spacyPipelineOptimisations=True}), lemmatised, and lowercased. POS is required (\texttt{usePOS=True}); noun detection uses spaCy POS tags (\texttt{useSpacyForConceptNounPOSdetection=True}). Concept columns are delimited by POS (\texttt{conceptColumnsDelimitByPOS=True}) using deterministic delimiter types \{VERB, ADP, POS tag, punctuation characters \texttt{;}, \texttt{:}, \texttt{.}, \texttt{?}, \texttt{!}, \texttt{.}\} and probabilistic delimiters between noun pairs (CCONJ/SCONJ, word types \texttt{is}, \texttt{are}, \texttt{,}, \texttt{(}). Consecutive nouns and hyphenated noun phrases are combined during pretraining (\texttt{pretrainCombineConsecutiveNouns=True}, \texttt{pretrainCombineHyphenatedNouns=True}).

\paragraph{Training regime.}
Training runs for 1 epoch (\texttt{numberEpochs=1}) with a maximum of 1000000 sequences (\texttt{trainMaxSequences<=1000000}), sequence length capped at 80 tokens (\texttt{maxSequenceLength=80}). \texttt{numSeedTokensInference=8,12,16} is retained for segment sizing. Single-sentence sequences are used by default (\texttt{multisentencePredictions=False}, \texttt{numSentencesPerSequence=1}). The database is stored at \texttt{/media/user/ssdpro/GIAANN/database/}.

\paragraph{Hardware.}
ASUS B760 Intel Chipset. ASUS NVIDIA GeForce RTX 5070 (12GB GDDR7). Intel i7 14700F. Patriot 64GB DDR4. Samsung 990 Pro PCIe 4.0 NVMe M.2 SSD (1TB).

\paragraph{Software.}
NVIDIA Driver 580.126.09. CUDA 12.8. Ubuntu 24.04 LTS. IceWM. NEdit. Codex IDE extension for Visual Studio Code \citep{openai2026codexIdeExtension}.

\paragraph{Configuration flags.}
\begin{itemize}
	\item Run mode: \texttt{useInference=False->True for train/eval}, \texttt{numSeedTokensInference=8,12,16}.
	\item Memory: \texttt{useGPUdense=True}, \texttt{useGPUsparse=True during train, useGPUsparse=False by default during eval}, \texttt{useGPUsparseStrict=True}.
	\item Tensors: \texttt{arrayIndexPropertiesEfficient=True}.
	\item Sequence observed columns: \texttt{trainSequenceObservedColumnsUseSequenceFeaturesOnly=True}, \texttt{trainSequenceObservedColumnsMatchSequenceWords=True}, \texttt{combineSparseUpdatesPerSequence=True}.
	\item SANI and segments: \texttt{useSANI=True}, \texttt{useSANIcolumns=False}, \texttt{useSANIfeatures=False}, \texttt{useSANIfeaturesAndColumns=True}, \texttt{arrayNumberOfSegmentsColumnDistance=floor(seed / 4) + 1}, \texttt{arrayNumberOfSegmentsFeatureDistance=ceil(seed / 2) + 1}, \texttt{arrayNumberOfSegments=arrayNumberOfSegmentsColumnDistance + arrayNumberOfSegmentsFeatureDistance}.
	\item Dendritic branches: \texttt{multipleDendriticBranches=True}, \texttt{numberOfDendriticBranches=2}, \texttt{randomlyAssignBranches=False} by default.
	\item Direct connections: \texttt{enforceDirectConnections=True} by default during train, \texttt{enforceDirectConnectionsSANI=True}, \texttt{enforceDirectConnectionsMinWordDistance=False}. \texttt{enforceDirectConnections=False} by default during eval of the training set.
	\item Beam search: \texttt{inferenceBeamSearch=False}, \texttt{inferenceBeamWidth=3}, \texttt {inferenceBeamDepth=3}.
	\item Inference timing (of segments): \texttt{inferenceUseNeuronFeaturePropertiesTime=True}, \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} by default during eval of the training set. \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} by default during eval of the test set.
	\item Inference activations: \texttt{inferenceConnectionsStrengthBoolean=True}, \texttt{inferenceSegmentActivationsBoolean=True} by default, \texttt{inferenceSegmentActivationsBooleanFeatureSegmentsOnly=True} by default, \texttt{inferenceSourceActivationsBoolean=True}.
	\item Inference prediction selection: \texttt{inferenceDeactivateNeuronsUponPrediction=True}, \texttt{inferenceDecrementActivations=False}.
	\item Drawing: \texttt{drawNetworkDuringTrain=False, drawNetworkDuringInference=False, drawDefault=True}, \texttt{drawSegments=False}, \texttt{drawBranches=False}, \texttt{drawRelationTypes=False}, \texttt{drawDelimiters=False}, \texttt{drawNetworkDuringTrainSave=True}.
	\item Spacy pipeline optimisations: \texttt{spacyPipelineOptimisations=False by default}, When true: \texttt{spacyPipelineSingleParse = False}, \texttt{spacyPipelineBatchSequences = True}, \texttt{spacyPipelineLightweightSentenceSegmentation = True}, \texttt{spacyPipelineMinimalComponents = True}.
\end{itemize}


\subsection{Metrics}
Top-1 accuracy has been calculated as a baseline performance metric.

\subsection{Results}

The accuracy measurements were taken with GIAANN Proto 1k. 

\subsubsection{Training-set Accuracy}

\paragraph{Tables}

\Cref{tab:wiki-train-top1-accuracy} contains Wikipedia top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input, and segment timing exact (``t-1 exact''). It includes `t-1 exact pred' (\texttt{inferenceUseNextTokenPredictionsOrTargetsToActivateNextColumnFeatures=True}), a non-standard top-1 accuracy report that is more sensitive to errors and can therefore better differentiate between ablations (it feeds current predictions rather than current targets to input).

\Cref{tab:wiki-train-top1-accuracy-segmentLayouts} contains Wikipedia top-1 training-set accuracy for different segment layouts and training sizes, with targets fed as input, and segment timing exact.

\Cref{tab:wiki-train-top1-accuracy-time} contains Wikipedia top-1 training-set accuracy for different sequentiality enforcement (\texttt{enforceSequentialActivation}), segment timing settings (\texttt{inferenceUseNeuronFeaturePropertiesTime/Exact}), and training sizes, with targets fed as input.

\Cref{tab:oscar-train-top1-accuracy} contains OSCAR-2201 top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input, and segment timing exact.

\Cref{tab:oscar-train-top1-accuracy-spacyPipelineOptimisations} likewise contains OSCAR-2201 top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input, and segment timing exact. It was trained with \url{spacyPipelineOptimisations=True} for efficiency.

\Cref{tab:oscar-train-top1-accuracy-useBenchmarkDefaultsFalse} contains OSCAR-2201 top-1 training-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input. The default time constraints used for benchmarking of the train-set(``t-1 exact'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} and \texttt{enforceSequentialActivation=True}. It also contains results for these non-default time constraints: \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} and \texttt{enforceSequentialActivation=True} (``top-1 bias''), and \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``top-1 none''). It was trained with \texttt{useBenchmarkDefaults=False} (\url{spacyPipelineOptimisations=True} and \texttt{pretrainConceptColumnsDelimitByPOSenforce=True}) for efficiency.

\Cref{tab:oscar-train-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions} contains OSCAR-2201 top-1 training-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input, and multiple sentence predictions. The default time constraints used for benchmarking of the train-set(``t-1 exact'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} and \texttt{enforceSequentialActivation=True}. It also contains results for these non-default time constraints: \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} and \texttt{enforceSequentialActivation=True} (``top-1 bias''), and \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``top-1 none''). It was trained with \texttt{multisentencePredictions=True} and \texttt{numSentencesPerSequence=3}, which generates longer sequences that are more suitable for higher segment counts. It was also trained with \texttt{useBenchmarkDefaults=False} (\url{spacyPipelineOptimisations=True} and \texttt{pretrainConceptColumnsDelimitByPOSenforce=True}) for efficiency.

\Cref{tab:oscar-train-top1-accuracy-spacyPipelineOptimisations-multipleDendriticBranches} contains OSCAR-2201 top-1 training-set accuracy for different training sizes, dendritic branches and time constraints, with targets fed as input. The default time constraints used for benchmarking of the train-set (``top-1 exact'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} and \texttt{enforceSequentialActivation=True}. It also contains results for these non-default time settings: \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} + \texttt{enforceSequentialActivation=False} (``t-1 bias''), \texttt{inferenceUseNeuronFeaturePropertiesTime=False} + \texttt{enforceSequentialActivation=False} (``t-1 none''), and activations constraints (with t-1 none): \texttt{inferenceSegmentActivationsBoolean=False} (``int f+c''). The default dendritic branch settings used for benchmarking are \texttt{multipleDendriticBranches=True}, \texttt{randomlyAssignBranches=False}, and \texttt{numberOfDendriticBranches=2} (``2 uniq''), where branches are uniquely assigned for repeated feature instances within sequences. It also contains results for these non-default dendritic branch settings: \texttt{randomlyAssignBranches=True}, and \texttt{numberOfDendriticBranches=5} (``5 rand''), \texttt{numberOfDendriticBranches=5} (``10 rand''). It was also trained with \texttt{spacyPipelineOptimisations=True} for efficiency.

The number of seed tokens, neuron column (c) segments, neuron feature (f) segments, and training sequences (along with their token top-1 accuracies), are provided in all tables.

\paragraph{Sequences}

The first 136 training-set sequences were used during inference for both the Wikipedia and OSCAR-2201 datasets.

A smaller subset of sequences ($\approx 90\%$) were processed by GIAANN during benchmarking, due to formal column segmentation parse rules \texttt{useBenchmarkDefaults: pretrainConceptColumnsDelimitByPOSenforce=False}. However it has been confirmed that top-1 accuracies do not significantly differ with/without these segmentation constraints. For seed length 12 for example, the train-set accuracy difference with \texttt{pretrainConceptColumnsDelimitByPOSenforce=True} is exactly -0.001 and -0.001 for 5000 and 100000 trained sequences respectively.

Wikipedia sentences are typically longer than OSCAR-2201 sentences, so multiple sentence predictions were only ever evaluated for OSCAR-2201.


\begin{table}[H]
\centering
\caption{Wikipedia Top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input.}
\label{tab:wiki-train-top1-accuracy}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 exact} & \textbf{t-1 exact pred} & \textbf{param} & \textbf{db size} \\
\midrule
8  & 3 & 5 & 5,000   & 0.985 & 0.968 & 3.0M & 0.2GB \\
8  & 3 & 5 & 100,000 & 0.984 & 0.935 & 57.0M & 3.3GB \\
8  & 3 & 5 & 500,000 & 0.973 & 0.819 & 271.2M & 15.1GB \\
12 & 4 & 7 & 5,000   & 0.982 & 0.983 & 3.2M & 0.2GB \\
12 & 4 & 7 & 100,000 & 0.982 & 0.960 & 62.0M & 3.6GB \\
12 & 4 & 7 & 500,000 & 0.977 & 0.945 & 296.8M & 16.8GB \\
16 & 5 & 9 & 5,000   & 0.997 & 0.999 & 3.5M & 0.2GB \\
16 & 5 & 9 & 100,000 & 0.996 & 0.996 & 70.2M & 4.2GB \\
16 & 5 & 9 & 500,000 & 0.993 & 0.979 & 336.6M & 19.5GB \\
\bottomrule
\end{tabular}
\end{table}

%\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 exact pred tok} & \textbf{t-1 exact pred seq} & \textbf{param} & \textbf{db size} \\
%8  & 3 & 5 & 5,000   & 0.978 & 0.975 \\	%orig manual calculations
%8  & 3 & 5 & 100,000 & 0.956 & 0.936 \\	%orig manual calculations
%8  & 3 & 5 & 540,000 & 0.882 & 0.817 \\	%orig manual calculations
%12 & 4 & 7 & 5,000   & 0.978 & 0.965 \\	%orig manual calculations
%12 & 4 & 7 & 100,000 & 0.978 & 0.954 \\	%orig manual calculations
%12 & 4 & 7 & 500,000 & 0.956 & 0.949 \\	%orig manual calculations
%(8) 12 & 0 & 7 & 5,000   & N/A & 0.975 \\
%(8) 12 & 0 & 7 & 100,000 & N/A & 0.946 \\
%12 & 4 & 0 & 5,000   & N/A & 0.000  \\
%12 & 4 & 0 & 100,000 & N/A & 0.000 \\



\begin{table}[H]
\centering
\caption{Wikipedia Top-1 training-set accuracy for different segment layouts and training sizes, with targets fed as input.}
\label{tab:wiki-train-top1-accuracy-segmentLayouts}
\begin{tabular}{rrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 exact} \\
\midrule
12 & 4 & 7 & 5,000   & 0.982 \\
12 & 4 & 7 & 100,000 & 0.982 \\
12 & 4 & 7 & 500,000 & 0.977 \\
(8) 12 & 0 & 7 & 5,000   & 0.974 \\
(8) 12 & 0 & 7 & 100,000 & 0.962 \\
12 & 4 & 0 & 5,000   & 0.000 \\
12 & 4 & 0 & 100,000 & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Wikipedia Top-1 training-set accuracy for different sequentiality enforcement, segment timing settings, and training sizes, with targets fed as input.}
\label{tab:wiki-train-top1-accuracy-time}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequentiality} & \textbf{timing} & \textbf{sequences} & \textbf{top-1} \\
\midrule
12 & 4 & 7 & enforced & exact & 5,000   & 0.982 \\
12 & 4 & 7 & enforced & exact & 100,000 & 0.982 \\
12 & 4 & 7 & enforced & exact & 500,000 & 0.977 \\
12 & 4 & 7 & enforced & biased & 5,000   & 0.983 \\
12 & 4 & 7 & enforced & biased & 100,000 & 0.979 \\
12 & 4 & 7 & enforced & biased & 500,000 & 0.959 \\
12 & 4 & 7 & enforced & none & 5,000   & 0.982 \\
12 & 4 & 7 & enforced & none & 100,000 & 0.976 \\
12 & 4 & 7 & enforced & none & 500,000 & 0.941 \\
12 & 4 & 7 & none & exact & 5,000   & 0.999 \\
12 & 4 & 7 & none & exact & 100,000 & 0.997 \\
12 & 4 & 7 & none & exact & 500,000 &  0.991 \\
12 & 4 & 7 & none & biased & 5,000   & 0.871 \\
12 & 4 & 7 & none & biased & 100,000 & 0.715 \\
12 & 4 & 7 & none & biased & 500,000 & 0.591 \\
12 & 4 & 7 & none & none & 5,000   & 0.980 \\
12 & 4 & 7 & none & none & 100,000 & 0.915 \\
12 & 4 & 7 & none & none & 500,000 & 0.807 \\
\bottomrule
\end{tabular}
\end{table}

%\begin{table}[H]
%\centering
%\caption{Wikipedia Top-1 training-set accuracy for different sequentiality enforcement, segment timing settings, and training sizes, with predictions fed as input.}
%\label{tab:wiki-train-top1-accuracy-time}
%\begin{tabular}{rrrrrrrr}
%\toprule
%\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequentiality} & %\textbf{timing} & \textbf{sequences} & \textbf{top-1} \\
%\midrule
%12 & 4 & 7 & enforced & exact & 5,000   & 0.983 \\
%12 & 4 & 7 & enforced & exact & 100,000 & 0.960 \\
%12 & 4 & 7 & enforced & exact & 500,000 &  \\
%12 & 4 & 7 & enforced & biased & 5,000   & 0.983  \\
%12 & 4 & 7 & enforced & biased & 100,000 & 0.951 \\
%12 & 4 & 7 & enforced & biased & 500,000 &  \\
%12 & 4 & 7 & enforced & none & 5,000   &  0.964 \\
%12 & 4 & 7 & enforced & none & 100,000 & 0.933 \\
%12 & 4 & 7 & enforced & none & 500,000 &  \\
%12 & 4 & 7 & none & exact & 5,000   &  1.000 \\
%12 & 4 & 7 & none & exact & 100,000 & 0.970 \\
%12 & 4 & 7 & none & exact & 500,000 &  \\
%12 & 4 & 7 & none & biased & 5,000   & 0.532 \\
%12 & 4 & 7 & none & biased & 100,000 & 0.253 \\
%12 & 4 & 7 & none & biased & 500,000 &  \\
%12 & 4 & 7 & none & none & 5,000   & 0.802 \\
%12 & 4 & 7 & none & none & 100,000 & 0.480 \\
%12 & 4 & 7 & none & none & 500,000 &  \\
%\bottomrule
%\end{tabular}
%\end{table}


\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input.}
\label{tab:oscar-train-top1-accuracy}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 exact} & \textbf{t-1 exact pred} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000    & 0.938 & 0.905 & 2.1M & 0.1GB \\
8 & 3 & 5 & 200,000  & 0.918 & 0.784 & 71.6M & 4.2GB \\
8 & 3 & 5 & 400,000  & 0.906 & 0.722 & 137.0M & 8.0GB \\
12 & 4 & 7 & 5,000   & 0.910 & 0.908 & 2.4M & 0.2GB \\
12 & 4 & 7 & 200,000 & 0.929 & 0.917 & 85.6M & 5.0GB \\
12 & 4 & 7 & 400,000 & 0.924 & 0.892 & 166.7M & 10.6GB \\
16 & 5 & 9 & 5,000   & 0.874 & 0.861 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000 & 0.890 & 0.871 & 103.0M & 6.2GB \\
16 & 5 & 9 & 400,000 & 0.888 & 0.844 & 198.4M & 11.8GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 training-set accuracy for different numbers of segments and training sizes, with targets/predictions fed as input.}
\label{tab:oscar-train-top1-accuracy-spacyPipelineOptimisations}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 exact} & \textbf{t-1 exact pred} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000   & 1.000 & 1.000 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000 & 0.987 & 0.947 & 76.6M & 4.5GB \\
12 & 4 & 7 & 5,000   & 1.000 & 1.000 & 2.5M & 0.2GB \\
12 & 4 & 7 & 200,000 & 1.000 & 1.000 & 88.6M & 5.3GB \\
16 & 5 & 9 & 5,000   & 1.000 & 1.000 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000 & 1.000 & 1.000 & 103.8M & 6.3GB \\
\bottomrule
\end{tabular}
\end{table}


\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 training-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input}
\label{tab:oscar-train-top1-accuracy-useBenchmarkDefaultsFalse}
\begin{tabular}{rrrrrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 exact} & \textbf{t-1 none} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.935 & 0.999 & 0.979 & 2.9M & 0.2GB \\
8 & 3 & 5 & 200,000   & 0.772 & 0.989 & 0.904 & 105.3M & 6.0GB \\
12 & 4 & 7 & 5,000    & 0.933 & 0.999 & 0.981 & 3.4M & 0.2GB \\
12 & 4 & 7 & 200,000  & 0.752 & 0.999 & 0.905 & 121.8M & 7.1GB \\
16 & 5 & 9 & 5,000    & 0.907 & 0.999 & 0.978 & 4.0M & 0.3GB \\
16 & 5 & 9 & 200,000  & 0.702 & 0.997 & 0.893 & 114.8M & 8.5GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 training-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input, and multiple sentence predictions}
\label{tab:oscar-train-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions}
\begin{tabular}{rrrrrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 exact} & \textbf{t-1 none} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.792 & 0.579 & 0.857 & 11.2M & 0.6GB \\
8 & 3 & 5 & 50,000    & 0.670 & 0.577 & 0.780 & 104.2M & 5.6GB \\
8 & 3 & 5 & 200,000   & 0.588 & 0.562 & 0.715 & 401.8M & 21.2GB \\
12 & 4 & 7 & 5,000    & 0.780 & 0.571 & 0.852 & 11.4M & 0.6GB \\
12 & 4 & 7 & 50,000   & 0.657 & 0.570 & 0.771 & 105.9M & 5.8GB \\
12 & 4 & 7 & 200,000  & 0.574 & 0.572 & 0.696 & 408.2M & 21.9GB \\
16 & 5 & 9 & 5,000    & 0.771 & 0.578 & 0.844 & 11.6M & 0.7GB \\
16 & 5 & 9 & 50,000   & 0.647 & 0.577 & 0.763 & 107.4M & 6.0GB \\
16 & 5 & 9 & 200,000  & 0.562 & 0.568 & 0.685 & 416.8M & 22.7GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 training-set accuracy for different training sizes, dendritic branches and time constraints, with targets fed as input.}
\label{tab:oscar-train-top1-accuracy-spacyPipelineOptimisations-multipleDendriticBranches}
\begin{tabular}{rrrrrrrrrrrrr}
\toprule
\textbf{s} & \textbf{c} & \textbf{f} & \textbf{sequences} & \textbf{dendrites} & \textbf{t1 bia} & \textbf{t1 exa} & \textbf{t1 non} & \textbf{int f+c} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 2 uniq & 0.933 & 1.000 & 0.983 & 0.950 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000   & 2 uniq & 0.746 & 0.987 & 0.901 & 0.859 & 76.6M & 4.5GB \\
8 & 3 & 5 & 5,000     & 5 rand & 0.933 & 1.000 & 0.984 & 0.948 & 2.1M & 0.1GB \\
8 & 3 & 5 & 200,000   & 5 rand & 0.748 & 0.989 & 0.904 & 0.865 & 76.8M & 4.6GB \\
8 & 3 & 5 & 5,000     & 10 rand & 0.933 & 0.999 & 0.984 & 0.949 & 2.1M & 0.1GB \\
8 & 3 & 5 & 200,000   & 10 rand & 0.748 & 0.987 & 0.904 & 0.863 & 76.8M & 4.6GB \\
\bottomrule
\end{tabular}
\end{table}

\FloatBarrier

\subsubsection{Test-set Accuracy}

\paragraph{Settings}

There is a tradeoff between recall/memorisation (train accuracy) and generalisation (test accuracy), and the optimum GIAANN parameters differ for both cases. Likewise, GIAANN uses word (vs subword) tokens and does not use word-vectors, which limits generalisation compared to contemporary gradient based models. GIAANN also relies on independent concept experts, which relates to robustness to hallucination.

Although GIAANN is designed to support prime concept feature name independence (by detecting all columns matching a prompt), this is computationally expensive as it requires bursting matching features across the entire trained network. Thus, test-set (prime) concept features must currently be present within the train-set. This limitation is strengthened by \texttt{enforceSequentialActivation} and \texttt{inferenceUseNeuronFeaturePropertiesTimeExact}.

Furthermore, the Wikipedia dataset cannot easily be used to perform a standardised NLP test-set accuracy measurement as it is divided into unique knowledgebase articles. Splitting the Wikipedia dataset into train/test portions would result in the train data being out of distribution with respect to the test data, especially when text is arranged/segmented into concepts (GIAANN).

Therefore, a number of changes were introduced for test-set accuracy measurement. 1. The \texttt{enforceSequentialActivation} and \texttt{inferenceUseNeuronFeaturePropertiesTimeExact} requirements were relaxed (i.e. disabled) during test-set inference. 2. the OSCAR-2201 dataset was exclusively used for test-set accuracy measurement.

\paragraph{Tables}

\Cref{tab:oscar-test-top1-accuracy} contains OSCAR-2201 top-1 test-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input. The default time constraints used for benchmarking of the test-set (``t-1 bias'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} and \texttt{enforceSequentialActivation=False}. It also contains results for these non-default time constraints: \texttt{enforceSequentialActivation=True} (``t-1 seq''),\texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} +  \texttt{enforceSequentialActivation=True} (``t-1 exact''), and \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``t-1 none'').

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations} contains OSCAR-2201 Top-1 test-set accuracy for different numbers of segments and training sizes, with segment timing biased, targets fed as input. It is an expanded version of the table across a greater selection of segment counts. It also contains results with \texttt{inferenceReportTokenAccuracyConstrainByColumn=True} (t-1 bias col). It was also trained with \texttt{spacyPipelineOptimisations=True} for efficiency.

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-trainsize} contains OSCAR-2201 Top-1 test-set accuracy for different training sizes, and different inference activations, with segment timing none (t-1 none), and targets fed as input. It is an expanded version of the table across a greater selection of train sizes. It was also trained with \texttt{spacyPipelineOptimisations=True} for efficiency.

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations} contains OSCAR-2201 top-1 test-set accuracy across different training sizes and inference activations, segment timing none (t-1 none), and targets fed as input. The different inference activation settings used were ``bool f+c'' (\texttt{inferenceSegmentActivationsBoolean=True}), ``int f+c'' (\texttt{inferenceSegmentActivationsBoolean=False}), and ``bool f'' (\texttt{inferenceSegmentActivationsBooleanFeatureSegmentsOnly=True} - default). It was also trained with \texttt{spacyPipelineOptimisations=True} for efficiency.

\Cref{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse} contains OSCAR-2201 top-1 test-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input.  The default time constraints used for benchmarking of the test-set (``t-1 bias'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} and \texttt{enforceSequentialActivation=False}. It also contains results for these non-default time constraints: \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} +  \texttt{enforceSequentialActivation=True} (``t-1 exact''), and \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``t-1 none''). It was trained with \texttt{useBenchmarkDefaults=False} (\url{spacyPipelineOptimisations=True} and \texttt{pretrainConceptColumnsDelimitByPOSenforce=True}) for efficiency and a greater use of the training data respectively.

\Cref{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions} contains OSCAR-2201 top-1 test-set accuracy for across different training sizes and time constraints, with multiple sentence predictions, and targets fed as input. It was trained with \texttt{multisentencePredictions=True} and \texttt{numSentencesPerSequence=3}, which generates longer sequences that are more suitable for higher segment counts with this dataset. It was also trained with \texttt{useBenchmarkDefaults=False} (\url{spacyPipelineOptimisations=True} and \texttt{pretrainConceptColumnsDelimitByPOSenforce=True}) for efficiency and a greater use of the training data respectively.

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference} contains another version of the previous table but with multiple sentence predictions used during test-set inference. Inference was effectively performed with a different test-set (\texttt{multisentencePredictions=True}) than usual for this train database.

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-multipleDendriticBranches} contains OSCAR-2201 top-1 test-set accuracy for different training sizes, dendritic branches and time constraints, with targets fed as input. The default time constraints used for benchmarking of the test-set (``t-1 bias'') are \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=False} and \texttt{enforceSequentialActivation=False}. It also contains results for these non-default time settings: \texttt{inferenceUseNeuronFeaturePropertiesTimeExact=True} + \texttt{enforceSequentialActivation=True} (``top-1 exact''), \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``top-1 none''), and activations constraints (with t-1 none): \texttt{inferenceSegmentActivationsBoolean=False} (``int f+c''). The default dendritic branch settings used for benchmarking are \texttt{multipleDendriticBranches=True}, \texttt{randomlyAssignBranches=False}, and \texttt{numberOfDendriticBranches=2} (``2 uniq''), where branches are uniquely assigned for repeated feature instances within sequences. It also contains results for these non-default dendritic branch settings: \texttt{randomlyAssignBranches=True}, and \texttt{numberOfDendriticBranches=5} (``5 rand''), \texttt{numberOfDendriticBranches=5} (``10 rand''). It was also trained with \texttt{spacyPipelineOptimisations=True} for efficiency.

149 test-set sequences were used during inference for the OSCAR-2201 dataset (taken from later in the dataset).

\paragraph{Discussion}

Notably, next-token prediction test accuracy in \Cref{tab:oscar-test-top1-accuracy}, \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations} and \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations} is higher with lower segment counts, indicating they are more likely to match (i.e. successfully continue) shorter subsequences versus longer subsequences. This trend is not consistent however, and is unobserved in \Cref{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions} and 
\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference}. It therefore appears to relate to the fact the test-set was being evaluted with \texttt{multisentencePredictions=False} in the first three tables. As OSCAR-2201 sentences are quite short relative to their Wikipedia counterparts, they do not always comprise the number of concept columns captured by their c segments for higher segment counts (seed lengths).

With both useSANIfeatures and useSANIfeaturesAndColumns, the first f segment is designated to accept predictions from all prior sequence tokens during train (when \texttt{SANIfeaturesLinkFirstSegmentToAllPriorTrainSeqTokens=True}), and the remaining f segments are designated to accept predictions from exactly f tokens prior to the predicted sequence token during train. This means the the last f segment is typically designated to accept predictions only from adjacent sequence tokens during train (assuming `f seg>=2' or \texttt{SANIfeaturesLinkFirstSegmentToAllPriorTrainSeqTokens=False}), which is used as a natural \texttt{enforceDirectConnectionsSANI} constraint.

Based on the results in \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations}, GIAANN Proto is not recognising many of the longer subsequences in the test-set when \texttt{multisentencePredictions=False}, and the majority of the measured predictive power is coming from detecting short adjacent 1-token subsequences. This predictive accuracy is coming from the last feature segment (\texttt{enforceDirectConnectionsSANI=True}). To illustrate this explicitly, the table includes entries for \texttt{c seg=0}, \texttt{f seg=1}, \texttt{useSANIfeatures=True}, generated using both a) \texttt{enforceDirectConnections=False} and \texttt{SANIfeaturesLinkFirstSegmentToAllPriorTrainSeqTokens=True} (its f segment captures predictions from all prior sequence tokens during train), and b) \texttt{enforceDirectConnections=True} and \texttt{SANIfeaturesLinkFirstSegmentToAllPriorTrainSeqTokens=False} (its f segment captures only the prior token during train).

\paragraph{Variations}

Next-token prediction accuracy reports are not restricted by column, meaning next token predictions of the correct word within the wrong the column are still reported as accurate. Top-1 test-accurcies have therefore also been reported when constrained by column, with \texttt{inferenceReportTokenAccuracyConstrainByColumn=True} (``top-1 col'') - see \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations}. The results indicate that there is no signficant degredation with this constraint enabled as the vast majority of correct predictions are occuring in the same column. With 5000 trained sequences for example, the test-set accuracy difference with \texttt{inferenceReportTokenAccuracyConstrainByColumn=True} is exactly 0.000, -0.001, and 0.000 for seed lengths 8, 12, and 16.

To improve evaluation efficency (speed), \texttt{spacyPipelineOptimisations=True} has been set in \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations}, \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-trainsize},  \Cref{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse}, \Cref{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions}, \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations}, \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference}, and \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-multipleDendriticBranches} as these optimisations have been confirmed to not significantly affect reported test-set accuracy. With 5000 trained sequences for example, the test-set accuracy difference with \texttt{spacyPipelineOptimisations=True} is exactly -0.002, for seed lengths 8, 12, and 16.

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations} and \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-trainsize} were generated using a different variation of the \texttt{enforceDirectConnectionsSANI} / \texttt{algorithmMatrixSANImethod="enforceActivationAcrossSegments} / \texttt{algorithmMatrixSANIenforceRequirement="enforceLastSegmentMustBeActive} code, which has since been deprecated. It has been confirmed that there is no significant differences between the test-set accuracies with these versions. With 5000 trained sequences for, the test-set accuracy differences with the deprecated version are exactly 0.000, 0.000, 0.000, -0.001, 0.000 for seed lengths  4, 6, 8, 12, 16. With 200000 trained sequences, the test-set accuracy differences with the deprecated version are exactly +0.001, 0.000, +0.001, 0.000, -0.002 for seed lengths 4, 6, 8, 12, 16.

Although the vast majority of inference evaluations were performed with \texttt{useGPUsparse=False}, some were performed with \texttt{useGPUsparse=True}. This resulted in a minor discrepancy in measured test accuracies. For example, a test that measured 0.026 with \texttt{useGPUsparse=False} was measured 0.030 with \texttt{useGPUsparse=True}. The vast majority of tables however were calculated with all rows using the same GPU settings, with \texttt{useGPUsparse=False} by default.

\paragraph{Ablations}

\Cref{tab:oscar-test-top1-accuracy} demonstrates that non-default segment time constraints performed better during test-set evaluation; \texttt{inferenceUseNeuronFeaturePropertiesTime=False} (``top-1 none''). 

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations} and \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference} demonstrate that non-default inference activations can perform better during test-set evaluation, and require further investigation; \texttt{inferenceSegmentActivationsBoolean=False} (``int f+c''). 

\Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations} and \Cref{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference} demonstrate that higher trained segment counts require longer evaluation sequences; \texttt{multisentencePredictions=True}.

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input.}
\label{tab:oscar-test-top1-accuracy}
\begin{tabular}{rrrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 seq} & \textbf{t-1 exa} & \textbf{t-1 none} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000   & 0.031 & 0.001 & 0.001 & 0.031 & 2.1M & 0.1GB \\
8 & 3 & 5 & 200,000 & 0.045 & 0.016 & 0.010 & 0.104 & 71.6M & 4.2GB \\
8 & 3 & 5 & 400,000 & 0.065 & 0.054 & 0.053 & 0.137 & 137.0M & 8.0GB \\
12 & 4 & 7 & 5,000   & 0.027 & 0.001 & 0.001 & 0.028 & 2.4M & 0.2GB \\
12 & 4 & 7 & 200,000 & 0.042 & 0.011 & 0.010 & 0.109 & 85.6M & 5.0GB \\
12 & 4 & 7 & 400,000 & 0.052 & 0.040 & 0.037 & 0.120 & 166.7M & 10.6GB \\
16 & 5 & 9 & 5,000   & 0.019 & 0.003 & 0.001 & 0.032 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000 & 0.042 & 0.023 & 0.023 & 0.126 & 103.0M & 6.2GB \\
16 & 5 & 9 & 400,000 & 0.045 & 0.024 & 0.023 & 0.135 & 198.4M & 11.8GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different numbers of segments and training sizes, with segment timing biased (t-1 bias), and targets fed as input.}
\label{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 bias col} & \textbf{param} & \textbf{db size} \\
\midrule
1 & 0 & adj 1 & 5,000   & 0.036 & 0.035 & 0.1M & 0.02GB \\
1 & 0 & adj 1 & 200,000 & 0.095 & 0.093 & 2.1M & 0.5GB \\
1 & 0 & all 1 & 5,000   & 0.011 & 0.011 & 0.9M & 0.1GB \\
1 & 0 & all 1 & 200,000 & 0.015 & 0.015 & 32.8M & 2.0GB \\
1 & 0 & 2 & 5,000   & 0.034 & 0.034 & 0.9M & 0.1GB \\
1 & 0 & 2 & 200,000 & 0.089 & 0.089 & 33.0M & 2.1GB \\
2 & 0 & 2 & 5,000   & 0.034 & 0.034 & 0.9M & 0.1GB \\
2 & 0 & 2 & 200,000 & 0.093 & 0.093 & 33.5M & 2.1GB \\
1 & 1 & 0 & 5,000   & 0.010 & 0.010 & 0.9M & 0.1GB \\
1 & 1 & 0 & 200,000 & 0.022 & 0.020 & 32.8M & 2.0GB \\
1 & 1 & 2 & 5,000   & 0.036 & 0.000 & 1.8M & 0.1GB \\
1 & 1 & 2 & 200,000 & 0.087 & 0.086 & 65.8M & 3.7GB \\
2 & 1 & 3 & 5,000   & 0.035 & 0.034 & 1.9M & 0.1GB \\
2 & 1 & 3 & 200,000 & 0.089 & 0.088 & 66.7M & 3.8GB \\
4 & 2 & 3 & 5,000   & 0.035 & 0.034 & 1.9M & 0.1GB \\
4 & 2 & 3 & 200,000 & 0.068 & 0.066 & 69.0M & 4.0GB \\
6 & 2 & 4 & 5,000   & 0.032 & 0.031 & 2.0M & 0.1GB \\
6 & 2 & 4 & 200,000 & 0.059 & 0.057 & 72.1M & 4.2GB \\
8 & 3 & 5 & 5,000   & 0.029 & 0.029 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000 & 0.051 & 0.049 & 76.6M & 4.5GB \\
12 & 4 & 7 & 5,000   & 0.025 & 0.024 & 2.5M & 0.2GB \\
12 & 4 & 7 & 200,000 & 0.060 & 0.060 & 88.6M & 5.3GB \\
16 & 5 & 9 & 5,000   & 0.017 & 0.017 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000 & 0.048 & 0.046 & 103.8M & 6.3GB \\
\bottomrule
\end{tabular}
\end{table}

%\begin{table}[H]
%\centering
%\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, with segment timing none (1-1 none), and targets fed as input.}
%\label{tab:oscar-test-top1-accuracy-trainsize}
%\begin{tabular}{rrrrrrr}
%\toprule
%\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{top-1 none} & \textbf{param} & \textbf{db size} \\
%\midrule
%8 & 3 & 5 & 5,000   & 0.031 & 2.5M & 0.2GB \\
%8 & 3 & 5 & 200,000 & 0.104 & 88.6M & 5.3GB \\
%8 & 3 & 5 & 400,000 & 0.137 & 88.6M & 5.3GB \\
%\bottomrule
%\end{tabular}
%\end{table}


\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, and different inference activations, with segment timing none (t-1 none), and targets fed as input.}
\label{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-trainsize}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{int f+c} & \textbf{bool f} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.029 & 0.029 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000   & 0.100 & 0.100 & 76.6M & 4.5GB \\
8 & 3 & 5 & 1,000,000 & 0.135 & 0.130 & 353.7M & 20.3GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, and different inference activations, segment timing none (t-1 none), and targets fed as input.}
\label{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations}
\begin{tabular}{rrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{bool f+c} & \textbf{int f+c} & \textbf{bool f} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.029 & 0.029 & 0.029 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000   & 0.091 & 0.100 & 0.100 & 76.6M & 4.5GB \\
12 & 4 & 7 & 5,000    & 0.028 & 0.029 & 0.031 & 2.5M & 0.2GB \\
12 & 4 & 7 & 200,000  & 0.110 & 0.116 & 0.117 & 88.6M & 5.3GB \\
16 & 5 & 9 & 5,000    & 0.023 & 0.023 & 0.024 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000  & 0.108 & 0.113 & 0.112 & 103.8M & 6.3GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different numbers of segments and training sizes, and different time constraints, with targets fed as input}
\label{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse}
\begin{tabular}{rrrrrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 exact} & \textbf{t-1 none} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.023 & 0.000 & 0.031 & 2.9M & 0.2GB \\
8 & 3 & 5 & 200,000   & 0.050 & 0.024 & 0.101 & 105.3M & 6.0GB \\
12 & 4 & 7 & 5,000    & 0.024 & 0.001 & 0.029 & 3.4M & 0.2GB \\
12 & 4 & 7 & 200,000  & 0.060 & 0.037 & 0.118 & 121.8M & 7.1GB \\
16 & 5 & 9 & 5,000    & 0.022 & 0.001 & 0.023 & 4.0M & 0.3GB \\
16 & 5 & 9 & 200,000  & 0.043 & 0.028 & 0.111 & 114.8M & 8.5GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, and different time constraints, with multiple sentence predictions, and targets fed as input.}
\label{tab:oscar-test-top1-accuracy-useBenchmarkDefaultsFalse-multisentencePredictions}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{t-1 bias} & \textbf{t-1 none} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.032 & 0.044 & 11.2M & 0.6GB \\
8 & 3 & 5 & 50,000    & 0.039 & 0.086 & 104.2M & 5.6GB \\
8 & 3 & 5 & 200,000   & 0.038 & 0.100 & 401.8M & 21.2GB \\
12 & 4 & 7 & 5,000    & 0.030 & 0.042 & 11.4M & 0.6GB \\
12 & 4 & 7 & 50,000   & 0.037 & 0.083 & 105.9M & 5.8GB \\
12 & 4 & 7 & 200,000  & 0.043 & 0.112 & 408.2M & 21.9GB \\
16 & 5 & 9 & 5,000    & 0.029 & 0.042 & 11.6M & 0.7GB \\
16 & 5 & 9 & 50,000   & 0.037 & 0.088 & 107.4M & 6.0GB \\
16 & 5 & 9 & 200,000  & 0.045 & 0.119 & 416.8M & 22.7GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, and different inference activations, with segment timing none (t-1 none), targets fed as input, and multiple sentence predictions (during eval only).}
\label{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-inferenceActivations-multisentencePredictionsInference}
\begin{tabular}{rrrrrrrrr}
\toprule
\textbf{seed} & \textbf{c seg} & \textbf{f seg} & \textbf{sequences} & \textbf{bool f+c} & \textbf{int f+c} & \textbf{bool f} & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 0.025 & 0.025 & 0.025 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000   & 0.074 & 0.084 & 0.083 & 76.6M & 4.5GB \\
12 & 4 & 7 & 5,000    & 0.027 & 0.028 & 0.028 & 2.5M & 0.2GB \\
12 & 4 & 7 & 200,000  & 0.089 & 0.098 & 0.094 & 88.6M & 5.3GB \\
16 & 5 & 9 & 5,000    & 0.027 & 0.028 & 0.028 & 2.9M & 0.2GB \\
16 & 5 & 9 & 200,000  & 0.095 & 0.099 & 0.099 & 103.8M & 6.3GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 Top-1 test-set accuracy for different training sizes, dendritic branches and time constraints, with targets fed as input.}
\label{tab:oscar-test-top1-accuracy-spacyPipelineOptimisations-multipleDendriticBranches}
\begin{tabular}{rrrrrrrrrrrrr}
\toprule
\textbf{s} & \textbf{c} & \textbf{f} & \textbf{sequences} & \textbf{dendrites} & \textbf{t1 bia} & \textbf{t1 exa} & \textbf{t1 non} & \textbf{int f+c}  & \textbf{param} & \textbf{db size} \\
\midrule
8 & 3 & 5 & 5,000     & 2 uniq & 0.029 & 0.000 & 0.029 & 0.029 & 2.2M & 0.1GB \\
8 & 3 & 5 & 200,000   & 2 uniq & 0.051 & 0.015 & 0.100 & 0.100 & 76.6M & 4.5GB \\
8 & 3 & 5 & 5,000     & 5 rand & 0.026 & 0.000 & 0.029 & 0.029 & 2.1M & 0.1GB \\	%seq: 0.000
8 & 3 & 5 & 200,000   & 5 rand & 0.047 & 0.015 & 0.101 & 0.090 & 76.8M & 4.6GB \\	%seq: 0.020
8 & 3 & 5 & 5,000     & 10 rand & 0.025 & 0.000 & 0.029 & 0.029 & 2.1M & 0.1GB \\
8 & 3 & 5 & 200,000   & 10 rand & 0.047 & 0.015 & 0.101 & 0.091 & 76.8M & 4.6GB \\
\bottomrule
\end{tabular}
\end{table}

\FloatBarrier

\subsection{Ablations}
Additional ablations to be conducted over:
\begin{itemize}
\item beam search depth/width,
\end{itemize}

\subsection{Performance}

The performance measurements were taken with GIAANN Proto 1l, after splitting the souce feature connections tensors per source feature.

Experimental optimisations ``optimisationGetFeatureConnectionsForSourceFeatureCache'' and ``optimisationObservedColumnsWriteMetadataCheck'' were disabled for all measurements.

\Cref{tab:oscar-performance-storeDatabaseInRamTrue} contains OSCAR-2201 performance measurements for different execution phases (train, eval training-set, eval test-set) and training sizes, with store database in RAM (``storeDatabaseInRam=True''). It also contains time measurements separated for the final write (train phase) or initial read (inference phase) of the entire database from/to CPU RAM (``t w/r''), where ``t tot = t proc + t w/r''. RAM measurements are calculated as peak memory allocated across execution.

\Cref{tab:oscar-performance-storeDatabaseInRamFalse} contains OSCAR-2201 performance measurements for different execution phases (train, eval training-set, eval test-set) and training sizes, without store database in RAM (``storeDatabaseInRam=False'').

\Cref{tab:oscar-performance-param} contains OSCAR-2201 model paramaters for different training sizes.

\begin{table}[H]
\centering
\caption{OSCAR-2201 performance measurements for different execution phases (train, eval training-set, eval test-set) and training sizes, with store database in RAM}
\label{tab:oscar-performance-storeDatabaseInRamTrue}
\begin{tabular}{rrrrrrrrrrrrr}
\toprule
\textbf{s} & \textbf{c} & \textbf{f} & \textbf{sequences} & \textbf{phase} & \textbf{t tot} & \textbf{t proc} & \textbf{t w/r} & \textbf{gpu ram} & \textbf{cpu ram} \\
\midrule
8 & 3 & 5 & 5,000       & train & 00:01:05 & 00:00:54 & 00:00:11 & 0.9GB & 1.9GB \\
8 & 3 & 5 & 200,000     & train & 01:30:38 & 01:26:02 & 00:04:36 & 1.9GB & 6.9GB \\
8 & 3 & 5 & 1,000,000   & train & 16:00:00 & 15:38:24 & 00:21:36 & 7.0GB & 20.6GB \\
8 & 3 & 5 & 5,000       & tr eval & 00:00:39 & 00:00:22 & 00:00:17 & 0.0GB & 1.6GB \\
8 & 3 & 5 & 200,000     & tr eval & 00:18:11 & 00:10:15 & 00:07:55 & 0.4GB & 7.1GB \\
8 & 3 & 5 & 1,000,000   & tr eval & 01:24:03 & 00:40:29 & 00:43:33 & 1.5GB & 23.3GB \\
8 & 3 & 5 & 5,000       & ts eval & 00:00:57 & 00:00:37 & 00:00:20 & 0.0GB & 1.6GB \\
8 & 3 & 5 & 200,000     & ts eval & 00:25:42 & 00:15:23 & 00:10:19 & 0.4GB & 7.2GB \\
8 & 3 & 5 & 1,000,000   & ts eval & 01:47:28 & 01:03:26 & 00:44:02 & 1.5GB & 20.6GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 performance measurements for different execution phases (train, eval training-set, eval test-set) and training sizes, without store database in RAM}
\label{tab:oscar-performance-storeDatabaseInRamFalse}
\begin{tabular}{rrrrrrrrrrr}
\toprule
\textbf{s} & \textbf{c} & \textbf{f} & \textbf{sequences} & \textbf{phase} & \textbf{t tot} & \textbf{gpu ram} & \textbf{cpu ram} \\
\midrule
8 & 3 & 5 & 5,000       & train & 00:01:19 & 0.9GB & 1.8GB \\
8 & 3 & 5 & 200,000     & train & 02:00:17 & 2.0GB & 2.0GB \\
8 & 3 & 5 & 5,000       & tr eval & 00:00:23 & 0.0GB & 1.5GB \\
8 & 3 & 5 & 200,000     & tr eval & 00:09:50 & 0.4GB & 2.2GB \\
8 & 3 & 5 & 5,000       & ts eval & 00:00:30 & 0.0GB & 1.5GB \\
8 & 3 & 5 & 200,000     & ts eval & 00:15:13 & 0.4GB & 2.2GB \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{OSCAR-2201 model paramaters for different training sizes}
\label{tab:oscar-performance-param}
\begin{tabular}{rrrrrrrr}
\toprule
\textbf{s} & \textbf{c} & \textbf{f} & \textbf{sequences} & \textbf{param} & \textbf{db size} & \textbf{columns} & \textbf{features} \\
\midrule
8 & 3 & 5 & 5,000       & 2.2M & 0.2GB & 9964 & 6263 \\
8 & 3 & 5 & 200,000     & 76.6M & 6.0GB & 203804 & 43636 \\
8 & 3 & 5 & 1,000,000   & 353.7M & 25.5GB & 795326 & 113756 \\
\bottomrule
\end{tabular}
\end{table}

\FloatBarrier

